---
output: github_document
---


<!-- README.md is generated from README.Rmd. Please edit that file -->


```{r setup, include = FALSE}
Sys.setlocale("LC_COLLATE", "en_US.UTF-8") # ensure common sorting envir
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
ver <- desc::desc_get_version(".")
ver <- paste0("https://img.shields.io/badge/Version-", ver,
              "-success.svg?style=flat&logo=github")
```


# The `pckpkg` package

<!-- badges: start -->
![GitHub version](`r ver`)
[![CRAN status](http://www.r-pkg.org/badges/version/pcapkg)](https://cran.r-project.org/package=pcapkg)
[![R-CMD-check](https://github.com/stufield/pcapkg/workflows/R-CMD-check/badge.svg)](https://github.com/stufield/pcapkg/actions)
[![](https://cranlogs.r-pkg.org/badges/grand-total/pcapkg)](https://cran.r-project.org/package=pcapkg)
[![test coverage](https://codecov.io/gh/stufield/pcapkg/branch/main/graph/badge.svg)](https://app.codecov.io/gh/stufield/pcapkg?branch=main)
[![lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![License: MIT](https://img.shields.io/badge/License-MIT-blue.svg)](https://choosealicense.com/licenses/mit/)
<!-- badges: end -->



## Overview

The `pcapkg` package contains the general functions necessary for the
unsupervised exploratory analysis of high-dimensional proteimic data using
Principal Component Analysis (PCA). The primary tools in this package are
dedicated to either the calculation of principal components, or the
visualization of loadings and projections of those components.


-----------------


## Installation

It is possible to install a specific version of `pcapkg` 

```{r install-git, eval = FALSE}
# current dev version
remotes::install_github("stufield/pcapkg")

# or a specific version
remotes::install_github("stufield/pcapkg@v0.0.1")
```


-----------------


## Usage

To load `pcapkg` simply make a call to `library()` as usual:

```{r, eval = FALSE}
library(pcapkg)
```


## Help summary of the package

```{r help, eval = FALSE}
library(help = pcapkg)
```


