---
title: "Introduction to pcapkg"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to pcapkg}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(pcapkg)
library(ggplot2)
library(patchwork)
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>"
)
```



The `pcapkg` package contains the general functions necessary for the
unsupervised exploratory analysis of high-dimensional proteomic data using
Principal Component Analysis (PCA). Primary tools related to calculation of
principal components and the visualization of rotation (loadings) and
projection spaces of these components.



---------------


## Useful functions in `pcapkg`

* `pca()`
  - perform principal component analysis (PCA) on a matrix
    of high-dimensional proteomic data.

* `plot_scree()`
  - generate a "scree"-plot of the top variances for a `pca` class object.

* `plot_rotation()`, `plot_projection()`
  - plot specific rotation and projection space objects

* `get_*()`
  - various "getter" functions to retreive features from the PCA.



-----------


## The PCA

Perform principal component analysis (PCA) on a 2-dim data matrix,
via the decomposition of the variance-covariance matrix (`SVD`).
Some modification of the standard `stats::prcomp()` is performed.
Data transformations (e.g. centering and/or scaling) can be performed
via its arguments.

```{r pca}
pr <- pca(pcapkg:::log10_ft(simdata))
pr
```

## Plotting

There are convenient plotting methods for both the
`rotation` and `projection` spaces.

### Rotation space

```{r rotation, fig.width = 6, fig.height = 5}
plot(pr, "r", identify = TRUE)
```

### Projection space

```{r projection, fig.width = 6, fig.height = 5}
plot(pr, "p", color = SiteId)                   # unquoted string!
plot(pr, "p", color = SiteId, id.labels = id)   # with labels
```

### Combine rotation and projection

```{r combo, fig.width = 10, fig.height = 5}
rot  <- plot(pr, "r", identify = TRUE)
proj <- plot(pr, "p", color = SiteId, id.labels = id) +
  theme(legend.position = "top")
rot + proj
```

### Corresponding `Scree` plots

```{r scree, fig.width = 10, fig.height = 5}
a <- plot_scree(pr, n = 10L)
b <- plot_scree(pr, n = 10L, type = "l")
a + b
```

